Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-11-07 introduction.tex
edits

In this paper, we propose a program logic for weakly-isolated
transactions along with automated verification support to allow
developers to verify the soundness of their applications, without
having to resort to low-level operational reasoning as they are forced
to do currently.  We develop a set of syntax-directed compositional
proof rules that enable the construction of correctness proofs for
transactional programs in the presence of a weakly-isolated
concurrency control mechanism.  Realizing that the proof burden
imposed by these rules may discourage applications programmers from
using them, we also present an inference procedure based on these
rules that automatically verifies the weakest isolation level for a
transaction that still ensures its invariants are maintained.  The key
to inference is a novel formulation of database state (represented as
sets of tuples) as a monad, and in which database computations are
interpreted as state transformers over these sets.  This
interpretation leads to an encoding of database computations amenable
for verification by off-the-shelf SMT solvers.  The paper makes the
following contributions:
%% We have realized these ideas as an embedded DSL within OCaml, that
%% supports common relational SQL operations (updates, selects, inserts,
%% deletes, joins, etc.).  Experimental results on real-world benchmarks
%% demonstrate the feasibility and value of automated verification for
%% weakly isolated transactions, an important advance in improving the
%% safety of realistic database computations.
\begin{enumerate}
  \item We analyze properties of weak isolation in the context of a
    DSL embedded in OCaml that treats SQL-based relational database
    operations (e.g., inserts, selects, deletes, updates, etc.) as
    computations over an abstract database state.
  \item We develop an operational semantics and a compositional
    rely-guarantee style proof system for this language capable of
    relating high-level application invariants to database state,
    parameterized by a weak isolation semantics that selectively
    exposes the visibility of these operations to other transactions.
  \item We devise an inference algorithm capable of discovering the
    weakest isolation level that is sound with respect to a
    transaction's high-level consistency requirements. The algorithm
    interprets database operations as state transformers expressed in
    a language amenable for translation into a decidable fragment of
    first-order logic, and is thus suitable for automated verification
    using off-the-shelf SMT solvers.
  \item We present details of an implementation along with an
    evaluation study on real database benchmarks that justify our
    approach, and demonstrate the utility of our inference mechanism.
\end{enumerate}
\noindent Our results provide the first (to the best of our knowledge)
formalization of weakly-isolated transactions, along with an
expressive and compositional proof automation framework capable of
verifying the safety of high-level consistency conditions attached to
these transactions.  Collectively, these contributions allow
weakly-isolated transactions to enjoy the same rigorous reasoning
capabilities as their strongly-isolated (serializable) counterparts.

The remainder of the paper is organized as follows. The next section
provides motivation and background on serializable and weakly-isolated
transactions. \S\ref{sec:opsem} presents an operational semantics for
a core language that supports weakly-isolated transactions,
parameterized over different isolation notions. \S\ref{sec:reasoning}
formalizes the proof system that we use to reason about program
invariants, and establishes the soundness of these rules with respect
to the semantics. \S\ref{sec:inference} describes the inference
algorithm, and the state transformer encoding.  We describe our
implementation in \S\ref{sec:implementation}, and provide case studies
and benchmark results in \S\ref{sec:case-studies}.  Related work is
given in \S\ref{sec:relatedwork}, and \S\ref{sec:conclusions} presents
conclusions.

>>>>

In this paper, we propose a program logic for weakly-isolated
transactions along with automated verification support to allow
developers to verify the soundness of their applications, without
having to resort to low-level operational reasoning as they are forced
to do currently.  We develop a set of syntax-directed compositional
proof rules that enable the construction of correctness proofs for
transactional programs in the presence of a weakly-isolated
concurrency control mechanism.  Realizing that the proof burden
imposed by these rules may discourage applications programmers from
using them, we also present an inference procedure that automatically
verifies the weakest isolation level for a transaction while ensuring
its invariants are maintained.  The key to inference is a novel
formulation of database state (represented as sets of tuples) as a
monad, and in which database computations are interpreted as state
transformers over these sets.  This interpretation leads to an
encoding of database computations amenable for verification by
off-the-shelf SMT solvers.  The paper makes the following
contributions:
%% We have realized these ideas as an embedded DSL within OCaml, that
%% supports common relational SQL operations (updates, selects, inserts,
%% deletes, joins, etc.).  Experimental results on real-world benchmarks
%% demonstrate the feasibility and value of automated verification for
%% weakly isolated transactions, an important advance in improving the
%% safety of realistic database computations.
\begin{enumerate}
  \item We analyze properties of weak isolation in the context of a
    DSL embedded in OCaml that treats SQL-based relational database
    operations (e.g., inserts, selects, deletes, updates, etc.) as
    computations over an abstract database state.
  \item We develop an operational semantics and a compositional
    rely-guarantee style proof system for this language capable of
    relating high-level application invariants to database state,
    parameterized by a weak isolation semantics that selectively
    exposes the visibility of these operations to other transactions.
  \item We devise an inference algorithm capable of discovering the
    weakest isolation level that is sound with respect to a
    transaction's high-level consistency requirements. The algorithm
    interprets database operations as state transformers expressed in
    a language amenable for translation into a decidable fragment of
    first-order logic, and is thus suitable for automated verification
    using off-the-shelf SMT solvers.
  \item We present details of an implementation along with an
    evaluation study on real database benchmarks that justify our
    approach, and demonstrate the utility of our inference mechanism.
\end{enumerate}
\noindent Our results provide the first formalization of
weakly-isolated transactions, along with an expressive and
compositional proof automation framework capable of verifying the
safety of high-level consistency conditions attached to these
transactions.  Collectively, these contributions allow weakly-isolated
transactions to enjoy the same rigorous reasoning capabilities as
their strongly-isolated (serializable) counterparts.

The remainder of the paper is organized as follows. The next section
provides motivation and background on serializable and weakly-isolated
transactions. \S\ref{sec:opsem} presents an operational semantics for
a core language that supports weakly-isolated transactions,
parameterized over different isolation notions. \S\ref{sec:reasoning}
formalizes the proof system that we use to reason about program
invariants, and establishes the soundness of these rules with respect
to the semantics. \S\ref{sec:inference} describes the inference
algorithm, and the state transformer encoding.  We describe our
implementation in \S\ref{sec:implementation}, and provide case studies
and benchmark results in \S\ref{sec:case-studies}.  Related work is
given in \S\ref{sec:relatedwork}, and \S\ref{sec:conclusions} concludes.

### 2017-11-07 motivation.tex
edits

We present our ideas in the context of a DSL embedded in OCaml that
manages an abstract database state, and allows it be manipulated via a
well-defined \C{SQL} interface. Arbitrary database computations can be
built around this interface, which can then be run as transactions
using the \C{atomically\_do} combinator provided by the DSL.

@@

Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundness of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes,
essentially yielding serial behaviors.  Under weak
isolation\footnote{Weak isolation does not violate atomicity as long
  as the witnessed effects are those of committed transactions},
however, interferences of various kinds are permitted, leading to
executions superficially similar to executions permitted by concurrent
(racy) programs~\cite{GHE15,HPQ+15}.  To illustrate, consider the
behavior of the \C{new\_order} transaction when executed under
\emph{Read Committed} (RC) isolation level, the default isolation
level in 8 of the 18 databases studied in~\cite{bailishotos}.
\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits before $T_1$. Nonetheless, both transactions read the same value for
  \C{d\_next\_o\_id} resulting in them  adding \C{Order} records
  with the same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\end{wrapfigure}
An executing RC transaction is isolated from \emph{dirty writes},
i.e., writes of uncommitted transactions, but is allowed to witness
the writes of concurrent transactions as soon as they are committed.
Thus, with two concurrent instances of the \C{new\_order} transaction
(call them $T_1$ and $T_2$), both concurrently placing new orders for
different customers under the same district (\C{d\_id}), RC isolation
allows the execution shown in Fig.~\ref{fig:new_order_exec}.

@@

The figure depicts an execution as a series of SQL operations. In the
execution, the \C{new\_order} instance $T_1$ (green) reads the
\C{d\_next\_o\_id} field of the district record for \C{d\_id}, but
before it increments the field, another \C{new\_order} instance
($T_2$) begins its execution and commits. Note that $T_2$ reads the
same \C{d\_next\_o\_id} value as $T_1$, and inserts new \C{Order} and
\C{Order\_line} records with their \C{o\_id} and \C{ol\_o\_id} fields
(resp.) equal to \C{d\_next\_o\_id}. $T_2$ also increments the
\C{d\_next\_o\_id} field, which $T_1$ has already acccessed. This is
allowed because reads typically do not obtain a mutually exclusive
lock on most databases. After $T_2$'s commit, $T_1$ resumes execution
and adds new \C{Order} and \C{Order\_line} fields with the same order
id as $T_1$. Thus, by the end of the execution, \C{Order\_line}
records inserted by $T_1$ and $T_2$ all bear the same order id. There
are also two \C{Order} records with the same district id (\C{d\_id})
and order id, none of whose \C{o\_ol\_cnt} reflects the actual number
of \C{Order\_line} records inserted with that order id.  This clearly
violates TPC-C's consistency condition.

Clearly, this example does not exhibit any of the anomalies that
\emph{characterize} RC isolation~\cite{berenson}\footnote{Berenson et
al~\cite{berenson} characterize isolation levels in terms of the
anomalies they \emph{admit}. For e.g., RC is characterized by
\emph{lost writes} because it admits the anomaly.}. For instance,
there are no \emph{lost writes} - both concurrent transactions' writes
are present in the final state of the database.  Program analyses that
aim to determine appropriate isolation by checking for possible
manifestations of RC-induced anomalies would fail to identify grounds
for promoting the isolation level of \C{new\_order} to something
stronger.  Yet, if we take the semantics of the application into
account, it is quite clear that RC is not an appropriate isolation
level for \C{new\_order}.

>>>>

We present our ideas in the context of a DSL embedded in OCaml that
manages an abstract database state that can be manipulated via a
well-defined \C{SQL} interface. Arbitrary database computations can be
built around this interface, which can then be run as transactions
using the \C{atomically\_do} combinator provided by the DSL.

@@

Because the aforementioned reasoning is reasonably simple to perform
manually, verifying the soundness of TPC-C's consistency conditions
would appear to be feasible.  Serializability aids the tractability of
verification by preventing any interference among concurrently
executing transactions while the \C{new\_order} transaction executes,
essentially yielding serial behaviors.  Under weak
isolation\footnote{Weak isolation does not violate atomicity as long
  as the witnessed effects are those of committed transactions},
however, interferences of various kinds are permitted, leading to
executions superficially similar to executions permitted by concurrent
(racy) programs~\cite{GHE15,HPQ+15}.  To illustrate, consider the
behavior of the \C{new\_order} transaction when executed under a
\emph{Read Committed} (RC) isolation level, the default isolation
level in 8 of the 18 databases studied in~\cite{bailishotos}.
\begin{wrapfigure}{l}{.4\textwidth}
\includegraphics[scale=0.45]{Figures/motiv-eg-1-b}
\caption{\small An RC execution involving two instances ($T_1$ and
  $T_2$) of the \C{new\_order} transaction depicted in
  Fig.~\ref{fig:new_order_code}. 
  Both instances read the \C{d\_id} \C{District} record concurrently,
  because neither transaction is committed when the reads are
  executed.  The subsequent operations are effectively sequentialized,
  since $T_2$ commits before $T_1$. Nonetheless, both transactions read the same value for
  \C{d\_next\_o\_id} resulting in them  adding \C{Order} records
  with the same ids, which in turn triggers a violation of TPC-C's
  consistency condition.}
\label{fig:new_order_exec}
\end{wrapfigure}
An executing RC transaction is isolated from \emph{dirty writes},
i.e., writes of uncommitted transactions, but is allowed to witness
the writes of concurrent transactions as soon as they are committed.
Thus, with two concurrent instances of the \C{new\_order} transaction
(call them $T_1$ and $T_2$), both concurrently placing new orders for
different customers under the same district (\C{d\_id}), RC isolation
allows the execution shown in Fig.~\ref{fig:new_order_exec}.

@@

The figure depicts an execution as a series of SQL operations. In the
execution, the \C{new\_order} instance $T_1$ (green) reads the
\C{d\_next\_o\_id} field of the district record for \C{d\_id}, but
before it increments the field, another \C{new\_order} instance
$T_2$ (red) begins its execution and commits. Note that $T_2$ reads the
same \C{d\_next\_o\_id} value as $T_1$, and inserts new \C{Order} and
\C{Order\_line} records with their \C{o\_id} and \C{ol\_o\_id} fields
(resp.) equal to \C{d\_next\_o\_id}. $T_2$ also increments the
\C{d\_next\_o\_id} field, which $T_1$ has already acccessed. This is
allowed because reads typically do not obtain a mutually exclusive
lock on most databases. After $T_2$'s commit, $T_1$ resumes execution
and adds new \C{Order} and \C{Order\_line} fields with the same order
id as $T_1$. Thus, at the end of the execution, \C{Order\_line}
records inserted by $T_1$ and $T_2$ all bear the same order id. There
are also two \C{Order} records with the same district id (\C{d\_id})
and order id, none of whose \C{o\_ol\_cnt} reflects the actual number
of \C{Order\_line} records inserted with that order id.  This clearly
violates TPC-C's consistency condition.

This example does not exhibit any of the anomalies that
\emph{characterize} RC isolation~\cite{berenson}\footnote{Berenson et
al~\cite{berenson} characterize isolation levels in terms of the
anomalies they \emph{admit}. For example, RC is characterized by
\emph{lost writes} because it admits the anomaly.}. For instance,
there are no \emph{lost writes} since both concurrent transactions' writes
are present in the final state of the database.  Program analyses that
aim to determine appropriate isolation by checking for possible
manifestations of RC-induced anomalies would fail to identify grounds
for promoting the isolation level of \C{new\_order} to something
stronger.  Yet, if we take the semantics of the application into
account, it is quite clear that RC is not an appropriate isolation
level for \C{new\_order}.

### 2017-11-07 reasoning.tex
edits

\noindent Thus, if $\rg{I,R}{c}{G,I}$ is a valid RG judgment, then (a)
every interleaved multi-step reduction of $c$ preserves the database
integrity constraint (consistency condition) $I$, and (b) the effect
that every transaction in $c$ has on the database state is captured by
$G$.
\noindent We can now assert the soundness of the RG judgments in
Fig.~\ref{fig:rg-rules} as follows\footnote{Full proofs for theorems
in this paper can be found in the accompanying tech report}:

>>>>

\noindent Thus, if $\rg{I,R}{c}{G,I}$ is a valid RG judgment, then (a)
every interleaved multi-step reduction of $c$ preserves the database
integrity constraint (consistency condition) $I$, and (b) the effect
that every transaction in $c$ has on the database state is captured by
$G$.
\noindent We can now assert the soundness of the RG judgments in
Fig.~\ref{fig:rg-rules} as follows\footnote{Full proofs for the major
  theorems and lemmas defined in this paper are available from
  ~\cite{KN+18_arxiv}.}:

### 2017-11-07 related-work.tex
edits

\vspace*{-6pt}
\section{Related Work}
\label{sec:relatedwork}

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, and the kinds of
dependencies that are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anomaly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
the declarative nature of our specifications make them suitable for
formal reasoning about program behavior. Sivaramakrishnan \emph{et
al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, making it difficult
to reason about a program as it builds the trace. Cerone \emph{et
al.}~\cite{gotsmanconcur15} specify isolation levels with atomic
visibility, but their specifications are also for complete traces.
Both the aforementioned specification frameworks use the vocabulary
introduced in~\cite{burckhardt14}. However, none of them are equipped
with a reasoning framework that can use such specifications to verify
programs under weak isolation.

Recent work by Crooks \emph{et al.}~\cite{CPA+17} also explores the
use of a state-based interpretation of isolation as we do, and share
our position that enables specification of weak isolation that are not
tied to implementation-specific artifacts.  However, they do not
consider verification (manual or automated) of client programs, and it
is not immediately apparent if their specification formalism is
amenable for use within a verification toolchain.  ~\cite{WB17}
presents a dynamic analysis for weak isolation that attempts to
discover weak isolation anomalies from SQL log files.  Their solution,
while capable of identifying database attacks due to the use of
incorrect (excessively weak) isolation levels, does not consider how
to verify application correctness, infer proper isolation levels, or
formally reason about the relationship between weak-isolation levels
and application invariants.

\vspace*{-4pt}
\paragraph{Reasoning under weak isolation.} In~\cite{feketessi}, Fekete
\emph{et al.} propose a theory to characterize non-serializable
executions that arise under {\sc si}. Fekete~\cite{fekete2005} also
proposes an algorithm that allocates either {\sc si} or {\sc ser}
isolation levels to transactions while guaranteeing
serializability. In~\cite{gotsmanpodc16}, Cerone \emph{et al.} improve
on Adya's {\sc si} specification and use it to derive a static
analysis that determines the safety of substituting {\sc si} with a
weaker variant called \iso{Parallel Snapshot Isolation}~\cite{psi}.
These efforts focus on establishing the equivalence of executions
between a pair of isolation levels, without taking application
invariants into account.  Bernstein \emph{et al.}~\cite{bern2000}
propose informal semantic conditions to ensure the satisfaction of
application invariants under weaker isolation levels.  All these
techniques are tailor-made for a finite set of well-understood
isolation levels (rooted in~\cite{berenson}) under a pre-defined store
consistency model.

\vspace*{-4pt}
\paragraph{Reasoning under weak consistency.} There have been several
recent proposals to reason about programs executing under weak
consistency~\cite{bailisvldb, alvarocalm, gotsmanpopl16,redblueatc,
redblueosdi, ecinec}. All of them assume a system model that offers a
choice between a \emph{coordination-free} weak consistency level
(\emph{e.g.}, eventual consistency~\cite{redblueosdi, redblueatc,
ecinec, alvarocalm, bailisvldb}) or causal
consistency~\cite{lbc16,gotsmanpopl16}). All these efforts involve
proving that atomic and fully isolated operations preserve application
invariants when executed under these consistency levels.  In contrast,
our focus in on reasoning in the presence of weakly-isolated
transactions under a strongly consistent store.  Gotsman \emph{et
al.}~\cite{gotsmanpopl16} adapt \iso{Parallel Snapshot Isolation} to a
transaction-less setting by interpreting it as a consistency level
that serializes writes to objects; a dedicated proof rule is developed
to help prove prove program invariants hold under this model. By
parameterizing our proof system over a gamut of weak isolation
specifications, we avoid the need to define a separate proof rule for
each new isolation level we may encounter.

@@

\vspace*{-4pt}
\paragraph{Inference.}  Prior work by Vafeiadis~\cite{Vaf10,Vaf10a}
describe \emph{action inference}, an inference procedure for computing
rely and guarantee relations in the context of RGSep~\cite{VP07}, an
integration of rely-guarantee and separation logic~\cite{Rey02} that
allows one to precisely reason about local and shared state of a
concurrent program. The ideas underlying action inference have been
used to prove memory safety, linearizability, shape invariant
inference, etc.  of fine-grained concurrent data structures.  While
our motivation is similar (automated inference of intermediate
assertions and local invariants), the context of study (transactions
vs. shared-memory concurrency), the objects being analyzed (relational
database tables vs. concurrent data structures), the properties being
verified (integrity constraints over relational tables vs. memory
safety, or linearizability of concurrent data structure operations)
and the analysis technique used to drive inference (state transformers
vs. abstract interpretation) are quite different.

>>>>

\section{Related Work}
\label{sec:relatedwork}

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, and the kinds of
dependencies that are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anomaly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
the declarative nature of our specifications make them suitable for
formal reasoning about program behavior. ~\cite{pldi15} specify
isolation levels declaratively as trace well-formedness conditions,
but their specifications implicitly assume a complete trace with only
committed transactions, making it difficult to reason about a program
as it builds the trace. ~\cite{gotsmanconcur15} specify isolation
levels with atomic visibility, but their specifications are also for
complete traces.  Both the aforementioned specification frameworks use
the vocabulary introduced in~\cite{burckhardt14}. However, none of
them are equipped with a reasoning framework that can use such
specifications to verify programs under weak isolation.

Recent work described in~\cite{CPA+17} also explores the
use of a state-based interpretation of isolation as we do, and 
like our approach, develops specifications of weak isolation that are not
tied to implementation-specific artifacts.  However, they do not
consider verification (manual or automated) of client programs, and it
is not immediately apparent if their specification formalism is
amenable for use within a verification toolchain.  ~\cite{WB17}
presents a dynamic analysis for weak isolation that attempts to
discover weak isolation anomalies from SQL log files.  Their solution,
while capable of identifying database attacks due to the use of
incorrect (excessively weak) isolation levels, does not consider how
to verify application correctness, infer proper isolation levels, or
formally reason about the relationship between weak-isolation levels
and application invariants.

\paragraph{Reasoning under weak isolation.} ~\cite{feketessi} propose
a theory to characterize non-serializable executions that arise under
{\sc si}. They also propose an algorithm that
allocates either {\sc si} or {\sc ser} isolation levels to
transactions while guaranteeing
serializability. ~\cite{gotsmanpodc16} improve
on Adya's {\sc si} specification and use it to derive a static
analysis that determines the safety of substituting {\sc si} with a
weaker variant called \iso{Parallel Snapshot Isolation}~\cite{psi}.
These efforts focus on establishing the equivalence of executions
between a pair of isolation levels, without taking application
invariants into account.  Bernstein \emph{et al.}~\cite{bern2000}
propose informal semantic conditions to ensure the satisfaction of
application invariants under weaker isolation levels.  All these
techniques are tailor-made for a finite set of well-understood
isolation levels (rooted in~\cite{berenson}).

\paragraph{Reasoning under weak consistency.} There have been several
recent proposals to reason about programs executing under weak
consistency~\cite{bailisvldb, alvarocalm, gotsmanpopl16,redblueatc,
redblueosdi, ecinec}. All of them assume a system model that offers a
choice between a \emph{coordination-free} weak consistency level
(\emph{e.g.}, eventual consistency~\cite{redblueosdi, redblueatc,
ecinec, alvarocalm, bailisvldb}) or causal
consistency~\cite{lbc16,gotsmanpopl16}). All these efforts involve
proving that atomic and fully isolated operations preserve application
invariants when executed under these consistency levels.  In contrast,
our focus in on reasoning in the presence of weakly-isolated
transactions under a strongly consistent store.  ~\cite{gotsmanpopl16}
adapt \iso{Parallel Snapshot Isolation} to a transaction-less setting
by interpreting it as a consistency level that serializes writes to
objects; a dedicated proof rule is developed to help prove prove
program invariants hold under this model. By parameterizing our proof
system over a gamut of weak isolation specifications, we avoid the
need to define a separate proof rule for each new isolation level we
may encounter.

@@

\paragraph{Inference.}  ~\cite{Vaf10,Vaf10a}
describe \emph{action inference}, an inference procedure for computing
rely and guarantee relations in the context of RGSep~\cite{VP07}, an
integration of rely-guarantee and separation logic~\cite{Rey02} that
allows one to precisely reason about local and shared state of a
concurrent program. The ideas underlying action inference have been
used to prove memory safety, linearizability, shape invariant
inference, etc.  of fine-grained concurrent data structures.  While
our motivation is similar (automated inference of intermediate
assertions and local invariants), the context of study (transactions
vs. shared-memory concurrency), the objects being analyzed (relational
database tables vs. concurrent data structures), the properties being
verified (integrity constraints over relational tables vs. memory
safety, or linearizability of concurrent data structure operations)
and the analysis technique used to drive inference (state transformers
vs. abstract interpretation) are quite different.
